---
title: "Regression Examples"
author: "Singh"

output:
  word_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r }
library(readxl)
Toluca_Example1 <- read_excel("C:/Users/Pradeep Singh/Documents/Professional/CSDA5110/Fall_2024/Week5/Toluca.xlsx")
View(Toluca_Example1)
```

## Scatter Plot



```{r }
library(ggplot2)
ggplot(
  data = Toluca_Example1,
  mapping = aes(x = lotsize, y = workhrs)) +
  geom_point() +
  geom_smooth(method = "lm")

```

## Summary Stats


```{r}
summary(Toluca_Example1)
```
## Regression Coefficients

```{r}
output.lm <- lm(workhrs ~ lotsize, Toluca_Example1)
output.lm
summary(output.lm)
residuals(output.lm)
fitted(output.lm)
fitted.values(output.lm)

```
##t-value 


##95% confidence intervals


```{r}
tval <- qt(1-0.05/2,23)
tval
round(confint(output.lm,level=0.95),3)

```
## Confidence and Prediction Intervals for complete data

```{r}
predict(output.lm, interval="confidence",level=0.95)
predict(output.lm,interval="prediction",level=0.95)

```
## Confidence and Prediction Intervals for Specific data points


```{r}
predict(output.lm,newdata=data.frame(lotsize=c(35,45,50,65,100,110)),interval="confidence",level=0.95)
predict(output.lm,newdata=data.frame(lotsize=c(35,45,50,65,100,110)),interval="prediction",level=0.95)

```
 ## ANOVA table
 
```{r}
anova(output.lm)

```
## Read Data Example 2
(S.J. Sheather, A Modern Approach to Regression with R)

```{r}
library(readxl)
anscombe <- read_excel("C:/Users/Pradeep Singh/Documents/Professional/CSDA5110/Fall_2024/Week5/anscombe.xlsx")
View(anscombe)
```

## Four Plots for four different pairs of data
```{r}
library(ggplot2)
ggplot(
  data = anscombe,
  mapping = aes(x = x1, y = y1)) +
  geom_point() +
  geom_smooth(method = "lm")

ggplot(
  data = anscombe,
  mapping = aes(x = x2, y = y2)) +
  geom_point() +
  geom_smooth(method = "lm")

ggplot(
  data = anscombe,
  mapping = aes(x = x3, y = y3)) +
  geom_point() +
  geom_smooth(method = "lm")

ggplot(
  data = anscombe,
  mapping = aes(x = x4, y = y4)) +
  geom_point() +
  geom_smooth(method = "lm")

```

##SLR for four pairs of data
```{r}
output1 <- lm(y1~x1, anscombe)
summary(output1)

output2 <- lm(y2~x2, anscombe)
summary(output2)

output3 <- lm(y3~x3, anscombe)
summary(output3)

output4 <- lm(y4~x4, anscombe)
summary(output4)

```

## Residual plots
```{r}
ggplot(
  data = output1,
  mapping = aes(x = x1, y = output1$residuals)) +
  geom_point()
  

ggplot(
  data = output2,
  mapping = aes(x = x2, y = output2$residuals)) +
  geom_point() 
  

ggplot(
  data = output3,
  mapping = aes(x = x3, y = output3$residuals)) +
  geom_point() 
  

ggplot(
  data = output4,
  mapping = aes(x = x4, y = output4$residuals)) +
  geom_point() 

```

## Normal Probablity plot and other plots

```{r}
plot(output1)
plot(output2)
plot(output3)
plot(output4)



```
## Multiple Linear Regression


```{r }
library(readxl)
Ital_Restau <- read_excel("C:/Users/Pradeep Singh/Documents/Professional/CSDA5110/Fall_2024/Week5/nyc.xlsx")

head(Ital_Restau)
data_restau <-Ital_Restau[,c(-1,-2,-7)]
data_restau

```

## MLR Estimation


```{r }
library(api2lm)
output.lm <- lm(Price ~ Food+Decor+Service+East, Ital_Restau)
output.lm
anova(output.lm)
summary(output.lm)

```
##Pairwise Correlation

```{r}
library(api2lm)
library(lmSubsets)
library(olsrr)
library(car)
 pairs( ~ Price+Food+Decor+Service+East, Ital_Restau)
 cor(data_restau)


```
## ANOVA Table

```{r}
anova(output.lm)
```



