---
title: "week-5"
author: "Rupa Shrestha"
date: "2024-09-26"
output: word_document
Student ID: 4266261
---

```{r}
library(readxl)
data_set <- read_excel("D:/Webster University/Assignments/Analytics Programming/week-5/Data_HW5.xlsx")
```
1) a) Obtain the scatter plot matrix and the correlation matrix. What information do these diagnostic aids provide here?
```{r}
#SOLUTION
library(ggplot2)
pairs(data_set)
cor(data_set)

# The plot shows the relationship between the dependent variable (Y) and the sum of two independent variables (X1 + X2). The points in the plot suggest a positive correlation between brand liking(Y) and with moisture content(X1) and sweetness (X2), meaning that as X1 + X2 increases, Y tends to increase as well. 
```
b. Fit the regression model to the data. State the estimated regression function. Interpret b1.
```{r}
#SOLUTION
# Fit the regression model
model <- lm(Y ~ X1 + X2, data = data_set)

# Display the summary of the model
summary(model)

# Extract coefficients
intercept <- coef(model)[1]
b1 <- coef(model)[2]
b2 <- coef(model)[3]

# Estimated regression function
#Y=β0+β1⋅X1+β2⋅X
print(paste("Estimated regression function: Y =", intercept, "+", b1, "* X1 +", b2, "* X2"))

# Interpret b1
print(paste("Interpretation of b1: For a one unit increase in X1, Y is expected to increase by", b1, "units, holding X2 constant."))

#The coefficient b1 represents the change in the degree of brand liking Y for each one-unit increase in moisture content X1, holding sweetness X2 constant
#For example, if b1 =3, it means that for every additional unit of moisture (assuming X1 increases by 1), the brand liking score Y increases by 3 points, regardless of the sweetness level X2.
```
c) Obtain the residuals and prepare boxplot. What information does it provide?
```{r}
#SOLUTION
residuals <- model$residuals
anova(model)
boxplot(residuals, main = "Residuals Boxplot", ylab = "Residuals")

# The boxplot shows the distribution of the residuals. The residuals are centered around zero and the spread is relatively small, suggesting a good model fit. There are a few outliers.
```
d) Plot the residuals against the fitted values, X1 and X2. Interpret them.
```{r}
#SOLUTION
# Plot residuals against fitted values
fitted_value <- model$fitted.values
X1 <- data_set$X1
X2 <- data_set$X2

plot(fitted_value, residuals, main="Residuals vs Fitted Values")
abline(h=0, col="red")
## Residuals vs Fitted Values: Residuals are randomly scattered around zero, suggesting no issues with the model's predictions across different fitted values.

# Plot residuals vs X1
plot(X1, residuals, main = "Residuals vs X1", xlab = "X1", ylab = "Residuals")
abline(h = 0, col = "blue")
## Residuals vs X1: Shows residuals scattered around zero with no clear pattern, indicating a good fit between X1 and the model.

# Plot residuals vs X2
plot(X2, residuals, main = "Residuals vs X2", xlab = "X2", ylab = "Residuals")
abline(h = 0, col = "green")
## Residuals vs X2: Shows residuals plotted against X2. The residuals are scattered randomly around zero, suggesting a good model fit.
```
e) Prepare a normal probability plot of the residuals and obtain the test statistics. State the results.
```{r}
#SOLUTION
#Normality Probability plot
plot(model)
qqnorm(residuals)
qqline(residuals, col = "red")
## The points fall along the diagonal line, the residuals are normally distributed.

# Shapiro-Wilk test for normality
shapiro_test <- shapiro.test(residuals)
shapiro_test
## Since p-value is greater than defalut value(0.05), so we fail to reject the null hypothesis. This indicates tha the residuals residuals do not significantly deviate from normality.
```
f) Test whether there is a regression relation, using α = .01. State the alternatives, decision rule, and conclusion. What does your test imply about β1 and β2?
```{r}
#SOLUTION
anova_result <- anova(model)
anova_result

```
```{r}
## Null Hypothesis (H0): There is no linear relationship between X1, X2, and Y (i.e., b1 = 0 and b2 = 0).

##Alternative Hypothesis (H1): At least one of X1 or X2 has a significant linear relationship with Y (i.e., at least one of b1 or b2 is not equal to 0).

##Decision Rule: Reject the null hypothesis if the p-value for either X1 or X2 is less than the significance level of 0.01.

##Conclusion: The p-values for both variables are significantly below 0.01, leading to the rejection of the null hypothesis. This indicates strong statistical evidence of a linear relationship with Y.

##Implications of b1 and b2: Both coefficients, b1 and b2, are significantly different from zero, indicating that X1 and X2 play a crucial role in predicting Y. Thus, both independent variables are important and should be included in the model.
```

g) Calculate the coefficient of multiple determination R2. How is it interpreted here?
```{r}
# SOLUTION
R_squared <- summary(model)$r.squared
print(paste("R² =", R_squared))
```
```{r}
## The high R-squared value indicates that 95.21% of the variability in Y can be accounted for by the independent variables X1 and X2.
## It indicates that the model explains substantial amount of the variance in brand liking, highlighting the effectiveness of moisture content and sweetness as predictors in regression model.
```

h) Obtain an interval estimate of E {Yx0} when X10 = 5 and X20 = 4. Use a 99 percent confidence coefficient. Interpret your interval estimate.
```{r}
#SOLUTION
# New data for prediction
new_data <- data.frame(X1 = 5, X2 = 4)

# 99% confidence interval
predict(model, newdata = new_data, interval = "confidence", level = 0.99)

```
```{r}
## The confidence inverval shows that for a product with moisture content of 5 and sweetness of 4, we can be 99% confident that the average degree of brand liking will be between apporx. 73.88 and 80.67. It helps to understand customer prefernce based on product characteristics.
```

i) Obtain a prediction interval for a new observation Y0(new) when X10 = 5 and X20. Use a 99 percent confidence coefficient.
```{r}
#SOLUTION
# Prediction interval
prediction_interval <- predict(model, newdata = new_data, interval = "prediction", level = 0.99)
print(prediction_interval)
```
```{r}
# We can be 99% confident that a new data of brand liking for a product with moisture content of 5 and sweetness of 4 will fall between 68.481 and 86.069.
```

2) The lengths of pregnancies in a small rural village are normally distributed with a mean of 265 days and a standard deviation of 15 days.
a. In what range would you expect to find the middle 50% of most pregnancies?
b. What proportion of pregnancies will be more than 275 days?
```{r}
# SOLUTION:

mean_pregnancy <- 265
sd_pregnancy <- 15

# a. Middle 50% (Q1 and Q3)
q1 <- qnorm(0.25, mean = mean_pregnancy, sd = sd_pregnancy)
q2 <- qnorm(0.75, mean=mean_pregnancy, sd=sd_pregnancy)
print(paste("Middle 50% range of pregnancies: ",q1 , q2))

# b. Proportion of pregnancies > 275 days
prob_more_275 <- 1-pnorm(275,mean = mean_pregnancy,sd=sd_pregnancy)
print(paste("Proportion of pregnancies more than 275 days: ", prob_more_275))


```
3) A certain car model has a mean gas mileage of 34 miles per gallon (mpg) with a standard deviation 5mpg. If a car is selected at random,
a) What is the probability that the mileage of the selected car is greater than 33 mpg?
b) What is the probability that the mileage of the car is between 33mpg and 35mpg?
```{r}
#SOLUTION
  
mean_gas_mileage <- 34
mileage_sd <-5

#a selected car mileage is greater than 33 mpg
prob_mileage_greater_than_33 <- 1-pnorm(33,mean=mean_gas_mileage,sd=mileage_sd)
print(paste("Probability that the mileage os selected car is greater than 33 mpg: ", prob_mileage_greater_than_33))

#b mileage of the car is between 33mpg and 35 mpg
prob_mileage_btn_33_35 <- pnorm(35, mean=mean_gas_mileage, sd=mileage_sd) - pnorm(33, mean=mean_gas_mileage, sd=mileage_sd)
print(paste("Probability of mileage between 33 and 35 is: ", prob_mileage_btn_33_35))

```
