---
title: "Chapter 7 Example 2"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

##Reading Data

```{r }
library(tidyverse)
library(dplyr)
library(ggplot2)
library(ggthemes)
library(readxl)
data_seatbelt <- read_excel("C:/Users/Pradeep Singh/OneDrive - Southeast Missouri State University/Teaching/Fall2024/CSDA5110/Fall_01/Data/SeatBelt_data.xlsx")
View(data_seatbelt)
head(data_seatbelt)

```

## Scatterplot


```{r }
ggplot(data = data_seatbelt, mapping = aes(x = Uninsured_Drivers, 
        y = Avg_Prem_Fluctuation_Creditscore) ) + geom_point()

```

##Scatterplot identifying Regions

```{r}

ggplot(data = data_seatbelt, mapping = aes(x = Uninsured_Drivers, 
                                           y = Avg_Prem_Fluctuation_Creditscore, color=Region) ) + geom_point()

```

## Scatterplot with trendline for each region
```{r}

ggplot(data = data_seatbelt, mapping = aes(x = Uninsured_Drivers, 
                  y = Prem_Fluctuation_CreditScore, color=Region) ) + geom_point() + geom_smooth(method = "lm")

```

##Scatterplot with single trendline
```{r}

ggplot(data = data_seatbelt, mapping = aes(x = Uninsured_Drivers, 
                                           y = Prem_Fluctuation_CreditScore) ) +
  geom_point(mapping = aes(color = Region)) +
  geom_smooth(method = "lm")

```

##Scatterplot with single trendline and regions identified by shape and color
```{r}

ggplot(data = data_seatbelt, mapping = aes(x = Uninsured_Drivers, 
                                           y = Avg_Prem_Fluctuation_Creditscore) ) +
  geom_point(mapping = aes(color = Region, shape=Region)) +
  geom_smooth(method = "lm")

```

##Scatterplot with labels
```{r}

ggplot(data = data_seatbelt, mapping = aes(x = Uninsured_Drivers, 
                                           y = Prem_Fluctuation_CreditScore) ) +
  geom_point(mapping = aes(color = Region, shape=Region)) +
  geom_smooth(method = "lm") +
  labs(   title = "Uninsured drivers vs Premium Fluctuation",
          subtitle = "Dimensions for Regions",
          x = "Uninsured drivers", y = "Premium Fluctuation",
          color = "Region", shape = "Region"
  ) +
  scale_color_colorblind()

```

##BarGraph
```{r}
ggplot(data_seatbelt, aes(x = Region)) +
  geom_bar()


```

##Bargraph - Pareto Chart
```{r}

ggplot(data_seatbelt, aes(x = fct_infreq(Region))) +
  geom_bar()

```

##Histogram

```{r}

ggplot(data_seatbelt, aes(x =Uninsured_Drivers )) +
  geom_histogram(binwidth = 10)

ggplot(data_seatbelt, aes(x =Avg_Prem_Fluctuation_Creditscore )) +
  geom_histogram(binwidth = 10)

```

##Density Plot
```{r}

ggplot(data_seatbelt, aes(x =Uninsured_Drivers)) +
  geom_density()

```

##Boxplot
```{r}

ggplot(data_seatbelt, aes(x = Region, y = Uninsured_Drivers)) +
  geom_boxplot()

```

##Density plots for each region for uninsured drivers
```{r}

ggplot(data_seatbelt, aes(x = Uninsured_Drivers, color = Region)) +
  geom_density(linewidth = 0.75)

```

##Summary Statistics
```{r}

median(data_seatbelt$Uninsured_Drivers)
mean(data_seatbelt$Uninsured_Drivers)
fivenum(data_seatbelt$Uninsured_Drivers)
var(data_seatbelt$Uninsured_Drivers)
sd(data_seatbelt$Uninsured_Drivers)

```

##Summary statistics for each level of categorical variable (Region)
```{r}

data_seatbelt %>% group_by(Region)  %>%
  summarise(mean_uninsured = mean(Uninsured_Drivers), std_dev_unsured= sd(Uninsured_Drivers),
            median_uninsured = median(Uninsured_Drivers),min_uninsured=min(Uninsured_Drivers),
            Q1=quantile(Uninsured_Drivers,probs=.25), Q3 =quantile(Uninsured_Drivers, probs=.75),
            max_uninsured=max(Uninsured_Drivers))

```

