---
title: "Chapter 11 & 12 Example"
output: word_document

---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

##Reading Data

```{r }
library(tidyverse)
library(dplyr)
library(ggplot2)
library(ggthemes)
library(readxl)
data_seatbelt <- read_excel("C:/Users/Pradeep Singh/Documents/Professional/CSDA5110/Fall_2024/Data/SeatBelt_data.xlsx")
View(data_seatbelt)
head(data_seatbelt)

```

## String function


```{r }
str(data_seatbelt)
str(data_seatbelt, strict.width ="cut")
data_seatbelt %>% str(strict.width = "cut")


```

## Functions

```{r}

length(data_seatbelt)
unique(data_seatbelt)
unique(data_seatbelt$Region)
length(unique(data_seatbelt$Region))
data_seatbelt$Region %>% unique() %>% length()


```


## Using the Apply() Family of Function

# The apply() function takes three arguments
# apply(X, MARGIN, FUN)
# x: an array or matrix
# MARGIN = 1: the manipulation is performed on rows
# MARGIN = 2: the manipulation is performed on columns, 
# MARGIN = c(1,2): the manipulation is performed on rows and columns
# FUN: tells which function to apply. Built functions like mean, median, sum, min, max and even user-defined functions can be applied


```{r}
class(data_seatbelt)
dim(data_seatbelt)
apply(data_seatbelt[,c(5,6)], 2 , median)

```

## Creating a matrix of numbers 1 through 10 with 5 rows and 5 columns
```{r}
m1 <- matrix(C<-(1:10),nrow=5, ncol=6)
m1
```

## using the apply() function to sum the matrices over all the rows
```{r}
rowsum_m1 <- apply(m1, 1, sum)
rowsum_m1

```

## using the apply() function to sum the matrices over all the columns
```{r}
colsum_m1 <- apply(m1, 2, sum)
colsum_m1

```
## Use lapply() to apply a function to a list

# first we create a list f movies
# Now we use the lapply() to convert the upper case to lower case by specifying the tolower function as the 2nd argument for lapply()\
# use the str() function to see each element
```{r}
movies <- c("SPIDERMAN","BATMAN","VERTIGO","CHINATOWN")
movies_lower <-lapply(movies, tolower)
str(movies_lower)
```

## Using the built in R data set "cars" we can see how lapply() and sapply() would work
# first we specify the min speed and stopping distance of the cars from the dataset using the min function as the 2nd argument
```{r}
data("cars")
dt <- cars
lmn_cars <- lapply(dt, min) # returns the results as a list
smn_cars <- sapply(dt, min) # returns the results as a vector

lmn_cars
class(lmn_cars)
smn_cars
class(smn_cars)

```

## Now we calculate the max speed and stopping distances of the cars
```{r}
lmxcars <- lapply(dt, max) # returns the results as a list
smxcars <- sapply(dt, mean)

lmn_cars
class(lmn_cars)
smxcars
class(smxcars)

```
## Now we create a function that calculates the average speed and stopping distances of the cars
```{r}
avg <- function(x) {  
  ( min(x) + max(x) ) / 2}
fcars <- sapply(dt, avg)
fcars

```
## Using Aggregate()
##Load the internal dataset chickwts that contains the weight and type of feed of a sample of chickens and assign it to a variable df.chic
```{r}
data("chickwts")
df.chic <- chickwts
head(df.chic)

names(df.chic)

dim(df.chic)

summary(df.chic)
```
## find the mean value for weight when grouped by feed using aggregate()
```{r}
group_mean <- aggregate(weight ~ feed, data = df.chic, mean)

# find the mean 
group_mean

# We can apply column manipulations by using the colnames() function and assign meaningful column names
colnames(group_mean) <- c("Group","Mean")
group_mean
```
##  use aggregate with more than one variable
```{r}
df.chickweight <- ChickWeight
names(ChickWeight)

aggregate(weight ~ Time + Diet, data = df.chickweight, mean)
```
## Aggregating with DPLYR

```{r}
# Find the mean weight depending on diet
group_by(df.chickweight, Diet) %>% summarise(mean = mean(weight))

# Aggregating the mean weight depending on Diet
group_by(df.chickweight, Time) %>% summarise(mean = mean(weight))

# Aggregating two variables
group_by(df.chickweight, Diet, Time) %>% summarise(mean = mean(weight))

# Aggregating on calculating two summaries
# Using aggregate
aggregate(weight ~ Diet, data = df.chickweight, FUN = function(x) c(mean = mean(x), n = length(x)))

# Using DPLYR
group_by(df.chickweight, Diet) %>% summarise(mean = mean(weight), n = length(weight))


# Aggregating on a subset of data 
# Using aggregate()
# The subset() function allows you to subset your data using logical operators
aggregate(weight ~ Diet, data = subset(df.chickweight, Diet!=2), mean)

# Using DPLYR
df.chickweight %>%
  filter(Diet != 1) %>%
  group_by(Diet) %>%
  summarise(mean = mean(weight))

# Using DPLYR to chain different functions using verbs
df.chickweight %>%
  filter(Diet != 1) %>%
  group_by(Diet) %>%
  summarise(mean = mean(weight)) %>%
  arrange(mean) # arrange() sorts from lowest to highest 

subset(df.chickweight, Diet == 1)
```

