---
title: "Probability Distributions"
output: word_document

---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Find Probabilities for Normal Distribution
###Blood Pressure Example
### Probability to the left


```{r }
library(stats)
Prob_X_less_than_100<-pnorm(100,120,20)
Prob_X_less_than_100

Prob_X_less_greater_133<- 1- pnorm(133,120,20)
Prob_X_less_greater_133

prob_between_100_133 <- (pnorm(133,120,20) - pnorm(100,120,20))
prob_between_100_133

```

## Finding X given area to the left



```{r}

Q1 <- qnorm(0.25, mean = 120, sd = 20)
Q1

```
## Finding X given area to the right

```{r}
X1 <- qnorm(0.90, mean = 120, sd = 20)
X1
```
## Find Probabilities for t- Distribution

```{r}
Prob_X_less_than_2<-pt(2,14)
Prob_X_less_than_2                     
```
## Find the value of t, when area to the right is given
```{r}
t1 <- qt(0.95, 14)
t1
```

## Binomial Distribution
Professor Example
```{r}
# P(X=6)
part1 <- dbinom(x=6,size=10, prob =.45)
part1

#P(X=0)
part2 <- dbinom(x=0,size=10, prob =.45)
part2

#P(X>=3)
part3 <- 1 - pbinom(q=2,size=10, prob =.45)
part3

```
## Teenage part-time job example
```{r}
#P(X>=3)
part4 <- 1 - pbinom(q=2,size=5, prob =.30)
part4
```
## Poisson Distribution
Typographical Error Example
```{r}
part5 <- dpois(x=3,lambda=.4)
part5
```
## McDonald's Example
```{r}
#P(X=6)
part6 <- dpois(x=6,lambda=10)
part6

#P(X< 6)
part7 <- ppois(q=5,lambda=10)
part7

#P(X>=6)
part8 <- 1- ppois(q=5,lambda=10)
part8
```
