---
title: "Week_3_Rupa_Shrestha"
output:
  word_document: default
  html_document: default
date: "2024-09-12"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
```{r}
#Name: Rupa Shrestha
#ID: 4266261
#Week3 Assignment
```

##Read the data and display the first and last six rows of the data.
```{r}
library(readxl)

housing_data=read_excel("D:/Webster University/Assignments/Analytics Programming/Week-3/house_selling_prices_FL.xlsx")

head(housing_data)
tail(housing_data)
```
##Rename ALL columns with FOUR letter words.
```{r}
library(stringr)

rename_four_letters_columns <- function(df) {
  column_names <- colnames(df)
  new_names <- sapply(column_names, function(name) {
    if (nchar(name) > 4) {
      paste0(str_sub(name, 1, 4))
    } else if (nchar(name) < 4) {
      paste0(str_pad(name, 4, side = "right", pad = "1"))
    } else {
      name
    }
  })
  colnames(df) <- new_names
  return(df)
}

df <- rename_four_letters_columns(housing_data)

colnames(df)
  
```
## Use the str() function to view the column names and data types for the table.
```{r}
str(df)
```
##Check the ‘type of ‘ variables Bedrooms and Baths. If they are not character, convert them to character
```{r}
class(housing_data$Bedrooms)
class(housing_data$Baths)

housing_data$Bedrooms <- as.character(housing_data$Bedrooms)
housing_data$Baths <- as.character(housing_data$Baths)

class(housing_data$Bedrooms)
class(housing_data$Baths)
```
##Display the value counts for Quadrant.
```{r}
value_count <- table(housing_data$Quadrant)
print(value_count)
```
## Add a column named PercentTax that contains the percentage of Tax out of the total price of the house.
```{r}
housing_data$PercentTax <- (housing_data$Taxes / housing_data$price) * 100
print(housing_data$PercentTax)
print(housing_data)
```
## Add a column named PercentSize that contains the percentage of house size out of the total lot size of the house.
```{r}
housing_data$PercentSize = (housing_data$size/housing_data$lot) * 100
print(housing_data$PercentSize)
print(housing_data)

```
## Use the apply() function to display the total PercentTax and total PercentSize.
```{r}
percent_size_matrix <- as.matrix(housing_data['PercentSize'])
percent_size_sum <- apply(percent_size_matrix,2,sum)

percent_tax_matrix <- as.matrix(housing_data['PercentTax'])
percent_tax_sum <- apply(percent_tax_matrix,2,sum)

print(percent_size_sum)
print(percent_tax_sum)
```
#. Group and summarize the data that includes newly added columns.
```{r}
library(dplyr)
summary_data <- housing_data %>% 
  group_by(Bedrooms) %>% 
  summarize(
    Total_Price = sum(price),
    Average_Percent_Tax=mean(PercentTax),
    Average_Percent_Size=mean(PercentSize), 
    .groups="drop"
    )

print(summary_data)
```
## Filter the data so that it only includes 3-Bedroom houses.
```{r}
filtered_housing_data = housing_data[housing_data$Bedrooms == 3,]
print(filtered_housing_data)
```



